TBD

### 2024-04-02 doc/paper.tex
Some writing about experiments

\usepackage[backgroundcolor=lightgray]{todonotes}
\usepackage{listings}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage[capitalize,noabbrev]{cleveref}
\usepackage{microtype}
\usepackage{mathtools}
\usepackage{forest}

@@

\newcommand{\expr}{\mathtt{expr}}
\newcommand{\FOtwo}{$\mathsf{FO}^{2}$}

@@

\todo[inline,caption={}]{
  \begin{itemize}
    \item check if I'm still using this notation in all these cases
    \item explain how this works for function calls too
  \end{itemize}
}

@@

\todo[inline,caption={}]{
  \begin{itemize}
    \item What notation am I using for FOL? Colon after every quantifier?
    \item (Similar to CNF, maybe QBF). Incomplete. Need (W)FOMC semantics. What
          is an interpretation? What is a model?
  \end{itemize}
}

@@

\paragraph{Remarks on comparing the model counts of the formulas with their
  corresponding OEIS sequences.}
Disregarding differences where the OEIS sequence starts with one or two numbers
that are not included in model counts (which is somewhat justified since we
start with domains of size one rather than zero), two differences remain:
\begin{itemize}
  \item on A62971, the OEIS sequence starts as $1, 2, 16,\dots$, but the model
        counts start as $1, 16,\dots$ (with the rest exactly the same);
  \item similarly, on A62971, the OEIS sequence starts as $32, 290,\dots$, but
        the model counts start as $2, 32, 290,\dots$ (with the rest exactly the
        same).
\end{itemize}

>>>>

\usepackage[backgroundcolor=lightgray]{todonotes}
\usepackage{listings}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage[capitalize,noabbrev]{cleveref}
\usepackage{microtype}
\usepackage{mathtools}
\usepackage{forest}
\usepackage{siunitx}
\usepackage{amsfonts}
\usepackage{pifont}
\usepackage[inline]{enumitem}

@@

\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

\newcommand{\expr}{\mathtt{expr}}
\newcommand{\Ctwo}{$\mathsf{C}^{2}$}
\newcommand{\FO}{$\mathsf{FO}$}
\newcommand{\FOtwo}{$\mathsf{FO}^{2}$}
\newcommand{\UFO}{$\mathsf{UFO}^{2} + \mathsf{EQ}$}

@@

\begin{itemize}
  \item Variables in uppercase
  \item Domains in uppercase Greek
  \item Constant and predicate symbols in lowercase
  \item Clauses written on separate lines are implicitly conjoined
  \item For weights, for any predicate symbol $p$, we write
        $w^{+}(p), w^{-}(p) \in \mathbb{R}$ for its positive and negative
        weights, respectively. All weights not explicitly acknowledged are
        assumed to be equal to one.
\end{itemize}

\todo[inline,caption={}]{
  \begin{itemize}
    \item check if I'm still using this notation in all these cases
    \item explain how this works for function calls too
    \item introduce WFOMC as a problem
  \end{itemize}
}

@@

\todo[inline,caption={}]{
  \begin{itemize}
    \item What notation am I using for FOL? Introduce equality cardinality
          constraints, counting quantifiers, etc.
    \item Need (W)FOMC semantics. What is an interpretation? What is a model?
    \item Should table caption go above or below the table?
    \item Define terms.
    \item Maybe say ``structure'' instead of interpretation?
    \item What's the right table heading capitalization for KR?
    \item Should I say ``sort'' instead of domain everywhere?
  \end{itemize}
}

\begin{table*}
  \centering
  \begin{tabular}{llllll}
    \toprule
    Logic & Sorts & Constants & Variables & Quantifiers & Additional atomic constraints\\
    \midrule
    \FO & one or more & \cmark & unlimited & $\forall$, $\exists$ & $s = t$\\
    \Ctwo & one & \xmark & two & $\forall$, $\exists$, $\exists^{= k}$, $\exists^{\le k}$, $\exists^{\ge k}$ & ---\\
    \UFO & one & \xmark & two & $\forall$ & $|p| = m$\\
    \bottomrule
  \end{tabular}
  \caption[]{A comparison of the three logics used in WFOMC in terms of:
    \begin{enumerate*}[label=(\roman*)]
      \item the number of sorts,
      \item support for constants,
      \item the maximum number of variables,
      \item allowed quantifiers, and
      \item atomic constructions in addition to those of the form $p(t_{1}, \dots, t_{n})$ for some predicate symbol $p$ and terms $t_{1}, \dots, t_{n}$.
    \end{enumerate*}
    Here:
    \begin{enumerate*}[label=(\roman*)]
      \item $k$ and $m$ are non-negative integers, the latter of which can depend on the domain size,
      \item $p$ is a predicate symbol, and
      \item $s$ and $t$ are terms.
    \end{enumerate*}
  }\label{tbl:logics}
\end{table*}

\paragraph{Three types of logics}
\begin{itemize}
  \item All three logics are function-free.
  \item Domains are always assumed to be finite.
  \item See \cref{tbl:logics} for a detailed comparison. \FO{} is used as the
        input format for \textsc{Crane} and \textsc{ForcLift}. \Ctwo{} is used
        in the literature on \textsc{FastWFOMC} and related methods (TODO:
        citations). \UFO{} is the input format supported by a privately-obtained
        version of \textsc{FastWFOMC}. Note that the publicly available version
        does not support any cardinality constraints.
  \item In the case of \textsc{Crane} and \textsc{ForcLift}, support for a
        formula as valid input does not imply that the algorithm will be able to
        compile the formula into a circuit or graph suitable for lifted model
        counting. However, it is known that \textsc{ForcLift} compilation is
        guaranteed to succeed on any \FO{} formula without constants and with at
        most two variables (TODO: citation).
\end{itemize}

@@

Comparing \textsc{Crane} and \textsc{FastWFOMC} on a larger set of benchmarks is
challenging because there is no automated way to translate a formula in \FO{} or
\Ctwo{} into \UFO{} (or even check if such an encoding is possible).

\paragraph{Benchmarks (probably for supplementary material).}
\begin{itemize}
  \item Functions
        \begin{itemize}
          \item In \Ctwo: $\forall X \in \Delta\text{.
                }\exists^{=1} Y \in \Delta\text{. }p(X, Y)$
          \item In \UFO:
                \begin{gather*}
                  \forall X, Y \in \Delta\text{. } s(X) \lor \neg p(X, Y)\\
                  |p| = |\Delta|
                \end{gather*}
          \item In \FO:
                \begin{gather*}
                  \forall X \in \Delta\text{. }\exists Y \in \Delta\text{. } p(X, Y)\\
                  \forall X, Y, Z \in \Delta\text{. } p(X, Y) \land p(X, Z) \Rightarrow Y = Z
                \end{gather*}
        \end{itemize}
  \item Permutations
        \begin{itemize}
          \item In \Ctwo:
                \begin{gather*}
                  \forall X \in \Delta\text{. }\exists^{=1} Y \in \Delta\text{. }p(X, Y)\\
                  \forall Y \in \Delta\text{. }\exists^{=1} X \in \Delta\text{. }p(X, Y)
                \end{gather*}
          \item In \UFO:
                \begin{gather*}
                  \forall X, Y \in \Delta\text{. } r(X) \lor \neg p(X, Y)\\
                  \forall X, Y \in \Delta\text{. } s(X) \lor \neg p(Y, X)\\
                  |p| = |\Delta|
                \end{gather*}
                with weights $w^{-}(r) = w^{-}(s) = -1$
          \item In \FO:
                \begin{gather*}
                  \forall X \in \Delta\text{. }\exists Y \in \Delta\text{. } p(X, Y)\\
                  \forall Y \in \Delta\text{. }\exists X \in \Delta\text{. } p(X, Y)\\
                  \forall X, Y, Z \in \Delta\text{. } p(X, Y) \land p(X, Z) \Rightarrow Y = Z\\
                  \forall X, Y, Z \in \Delta\text{. } p(X, Y) \land p(Z, Y) \Rightarrow X = Z
                \end{gather*}
        \end{itemize}
\end{itemize}

\paragraph{Setup.}
\begin{itemize}
  \item The experiments were run on an AMD~Ryzen~7~5800H processor with
        \SI{16}{\gibi\byte} of memory and Arch Linux~6.8.2-arch2-1 operating
        system. \textsc{FastWFOMC} was run using Python~3.8.19 with
        Python-FLINT~0.5.0.
\end{itemize}

\begin{figure}
  \centering
  \input{plot}
  \caption{The runtime data of WFOMC algorithms on bijection- and
    function-counting problems on domains of sizes
    $1, 2, \dots, 35$}\label{fig:plot}
\end{figure}

\paragraph{Results.}
\begin{itemize}
  \item the runtime appears constant because compilation time dominates
        inference time for small domain sizes. Indeed, the maximum inference
        time of both \textsc{Crane-BFS} and \textsc{Crane-Greedy} is only
        \SI{4}{\milli\second}.
  \item describe \cref{fig:plot}
  \item On bijection counting, \textsc{FastWFOMC} takes about \SI{124}{\second}
        on a domain of size 35, whereas \textsc{Crane-BFS} scales up to
        \num{e4} in \SI{34}{\second}
  \item On function counting, \textsc{FastWFOMC} takes about \SI{78}{\second} on
        a domain of size 35, whereas both \textsc{Crane-BFS} and
        \textsc{Crane-Greedy} scale up to \num{3e5} in \SI{41}{\second}.
  \item Note that, both model counts are huge (TODO: add details)
\end{itemize}
% ** bijections max time: 123.870
% *** Crane-BFS scales up to 10k in 34 s
% **** answer = 10k! = a number with 35660 digits
% ** functions max time: 78.330
% *** both Crane-BFS and Crane-Greedy scale up to 300k in 41 s
% **** the count of this is 300k^300k is a number with over 1 million digits

\todo[inline,caption={}]{
  Some reproducibility requirements to keep in mind:
  \begin{itemize}
    \item A motivation is given for why the experiments are conducted on the
          selected datasets.
    \item All novel datasets introduced in this paper are included in a data
          appendix.
    \item All datasets drawn from the existing literature (potentially including
          authors’ own previously published work) are accompanied by appropriate
          citations. (mention the counting quantifier paper and my KR paper)
    \item All source code implementing new methods have comments detailing the
          implementation, with references to the paper where each step comes
          from.
    \item This paper formally describes evaluation metrics used and explains the
          motivation for choosing these metrics.
    \item This paper states the number of algorithm runs used to compute each
          reported result.
  \end{itemize}
}
